Efficiently Handling Guards in the low level design of RPython's tracing JIT

### 2012-07-18 talk/vmil2012/paper.tex
start writing about guards <-> bridges in the backend

   * tracing and attaching bridges and throwing away resume data
   * compiling bridges

% section Resume Data (end)

@@

Code generation consists of two passes over the lists of instructions, a
backwards pass to calculate live ranges of IR-level variables and a forward one
to emit the instructions. During the forward pass IR-level variables are
assigned to registers and stack locations by the register allocator according
to the requirements of the to be emitted instructions. Eviction/spilling is
performed based on the live range information collected in the first pass. Each
IR instruction is transformed into one or more machine level instructions that
implement the required semantics, operations withouth side effects whose result
is not used are not emitted. Guards instructions are transformed into fast
checks at the machine code level that verify the corresponding condition.  In
cases the value being checked by the guard is not used anywhere else the guard
and the operation producing the value can merged, reducing even more the
overhead of the guard.

@@

>>>>

   * tracing and attaching bridges and throwing away resume data
   * compiling bridges
\bivab{mention that the failargs also go into the bridge}
% section Resume Data (end)

@@

Code generation consists of two passes over the lists of instructions, a
backwards pass to calculate live ranges of IR-level variables and a forward one
to emit the instructions. During the forward pass IR-level variables are
assigned to registers and stack locations by the register allocator according
to the requirements of the to be emitted instructions. Eviction/spilling is
performed based on the live range information collected in the first pass. Each
IR instruction is transformed into one or more machine level instructions that
implement the required semantics, operations withouth side effects whose result
is not used are not emitted. Guards instructions are transformed into fast
checks at the machine code level that verify the corresponding condition.  In
cases the value being checked by the guard is not used anywhere else the guard
and the operation producing the value can merged, reducing even more the
overhead of the guard. \bivab{example for this}

@@

As explained in previous sections, when a specific guard has failed often enogh
a new trace, refered to as bridge, starting from this guard is recorded and
compile. The goal for the execution of bridges, that become a part of the
common path is to favor the performance while staying on the compiled trace.
This means we want wo avoid switching back and forth to the frontend when a
guard corresponding to the bridge fails before we can execute the brigde.
Instead we want to have as little as possible overhead when switching from the
loop path to the bridge.

The process of compiling a bridge is very similar to compiling a loop,
instructions and guards are processed in the same way as described above. The
main difference is the setup phase, when compiling a trace we start with a
clean slate, whilst the compilation of a bridge starts with state as it was
when compiling the guard. To restore the state needed compile the bridge we use
the encoded representation created for the guard to rebuild the bindings from
IR-variables to stack locations and registers used in the register allocator.

Once the bridge has been compiled the trampoline method stub is redirected to
the code of the bridge. In future if the guard fails again it jumps to the
trampoline and then jumps to the code compiled for the bridge, having alsmos no
overhead compared to the execution of the original trace, behaving mainly as
two paths in a conditional block.
